Results

### 2018-11-15 Manuscript_draft/Manuscript_draft.Rmd
little by little

>>>>

Traditionally heterospecific pollen effect has focused its attention on different pollen donors as a main driver of different effect. However in this article we want to emphasize that this is true for the cases that the species are higly close related where pollen recognition can take place (eg hybridization) but not when this pollen is from less closely related species which the main driver of effect is determined by the reproductive biology of the female part of the plant(compatibility system, stigma type, stigma area and number of ovules).

### 2018-11-15 Manuscript_draft/Results.Rmd
little by little

Eq.1  $$HpEffect_{i}=Mean(CrosPollinationSeeds)_{i}-Mean(TreatmentSeeds_{ij})$$

@@

I plot now the pairwise distances on the x axis and the effect of heterospecific pollen on the y axis. The greater the effect of heterospecific pollen the greater the number on the y axis.

```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Figure 1. Scatterplot of the evolutive distances (x axis) and the mean scaled effect per species (ITS markers). In other words, each point is defined by the phylogenetic distnace between a pair of species (x axis) and the effect of heterospecifc pollen which considers also an interaction between the same pair of species (donor and recipient)(y axis)"}
its_distance <- melt(evo_distance_its)
scaled_effect <- melt(matrix_scale_effect)
scaled_effect$Species=as.character(scaled_effect$Species)
scaled_effect$col_focal[scaled_effect$Species==c("BROL")] <- "red"
scaled_effect$col_focal[scaled_effect$Species==c("BRRA")] <- "red"
scaled_effect$col_focal[scaled_effect$Species==c("SIAL")] <- "red"
scaled_effect$col_focal[scaled_effect$Species==c("ERSA")] <- "red"
scaled_effect$col_focal[scaled_effect$Species==c("SOME")] <- "blue"
scaled_effect$col_focal[scaled_effect$Species==c("SOLY")] <- "blue"
scaled_effect$col_focal[scaled_effect$Species==c("PEIN")] <- "blue"
scaled_effect$col_focal[scaled_effect$Species==c("CAAN")] <- "blue"
scaled_effect$col_focal[scaled_effect$Species==c("IPAQ")] <- "green"
scaled_effect$col_focal[scaled_effect$Species==c("IPPU")] <- "green"
plot(its_distance$value, scaled_effect$value, main="Hp effect & phylogenetic distance", 
     xlab="Pairwise phylogenetic distances", ylab="Hp effect (Scaled seed set)", pch=19, col=scaled_effect$col_focal)
draw.circle(0.265,0.22,radius=0.045,nv=100,border="red",col=NA,lty=1,density=NULL,angle=45,lwd=1)
draw.circle(0.245,1.8,radius=0.07,nv=100,border="blue",col=NA,lty=1,density=NULL,angle=45,lwd=1)
focal <- cbind(scaled_effect,its_distance)
focal <- data.frame(focal, stringsAsFactors = F)
#str(focal)
focal_bra <- filter(focal, Species==c("BROL","BRRA","ERSA","SIAL"))
focal_sol <- filter(focal, Species==c("SOME","SOLY","PEIN","CAAN"))
focal_con <- filter(focal, Species==c("IPAQ","IPPU"))
model_bra <- lm(focal_bra$value ~ focal_bra$value.1)
#summary(model_brra)
abline(model_bra, col="red")
model_sol <- lm(focal_sol$value ~ focal_sol$value.1)
#summary(model_sol)
abline(model_sol, col="blue")
model_con <- lm(focal_con$value ~ focal_con$value.1)
#summary(model_con)
abline(model_con, col="black")
```

Overall, it seems that with phylogenetic distance, heterospecific pollen effect increases for the three families. From Mantel analysis we can conclude that we have a significant correlation between phylogenetic distance and heterospecific pollen effect with r values between 0.25 and 0.3. However, for procustes test we were not able to find a significant correlation between these two despite the procustes correlation gave higher values. In Figure 1 we can also see these general pattern of a higher effect with phylogenetic distance with the different regressions lines per family. I have done simple linear models, probably I should consider other alternatives that fit better the cloud of points. In Figure 1 the two coulored circles are there for a reason, once we consider just the crosses between the "far related species" we see a general pattern for solanaceae with almost all the values above 0.5. Solanaceae curiously belongs to the group with gametophytic incompatibility which is predicted to suffer more from heterospecific pollen. Moreover, for Brassicaceae (red points) again for the far related crosses we find kind of two groups with high effect and no effect. The no effect correspond to the species without incompatibility (over 50% of selfing) and the high effect to the species with strong incompatibility (0% of selfing), again this confirms theory.

>>>>

Eq.1  $$HpEffect_{i}=Mean(CrosPollinationSeeds_{i})-Mean(TreatmentSeeds_{ij})$$

The number of possible combinations with 10 species (multiplied by two because is in both directions the treatments A to B and B to A):

$$\frac{10!}{2!*(8!)}*2=90 $$

@@

Overall, it seems that with phylogenetic distance, heterospecific pollen effect increases for the three families. From Mantel analysis we can conclude that we have a significant correlation between phylogenetic distance and heterospecific pollen effect with r values between 0.25 and 0.3. However, for procustes test we were not able to find a significant correlation between these two despite procustes correlation gave higher values. This positive correlation should be further explored. Maybe a scatterplot of the phylogenetic distance and the heterospecific pollen effect could help to understand this correlation.
```{r, echo=FALSE, message=FALSE, warning=FALSE}

plot(focal$hp_effect ~ focal$its_distance, main="", xlim=c(0.05,0.6),
     xlab="Evolutive distance", ylab="Hp effect", pch=19, col=focal$col_focal)

legend(x=0.04,y=3.2,legend=c("Brassicaceae","Convolvulaceae","Solanaceae"),pch=(c(16,16,16)), 
       col = c("red","black","blue"), bty="n",cex=0.6,pt.cex=0.6,xpd=TRUE)

```
Figure 1. Scatterplot of the evolutive distances (x axis) and the mean scaled effect per species (ITS markers). In other words, each point is defined by the phylogenetic distance between a pair of species (x axis) and the effect of heterospecifc pollen which considers also an interaction between the same pair of species (donor and recipient)(y axis)
\newpage

Although is not very elegant, I  had the curiosity of how a regression line will fit these 3 different families. Here I show how they wit the colud of points per family *lm(HpEffect~its_distance)*.

```{r, echo=FALSE, message=FALSE, warning=FALSE}

plot(focal$hp_effect ~ focal$its_distance, main="", xlim=c(0.05,0.6),
     xlab="Evolutive distance", ylab="Hp effect", pch=19, col=focal$col_focal)

legend(x=0.04,y=3.2,legend=c("Brassicaceae","Convolvulaceae","Solanaceae"),pch=(c(16,16,16)), 
       col = c("red","black","blue"), bty="n",cex=0.6,pt.cex=0.6,xpd=TRUE)

model_bra <- lm(focal_bra$value ~ focal_bra$value.1)
#summary(model_brra)
abline(model_bra, col="red")
model_sol <- lm(focal_sol$value ~ focal_sol$value.1)
#summary(model_sol)
abline(model_sol, col="blue")
model_con <- lm(focal_con$value ~ focal_con$value.1)
#summary(model_con)
abline(model_con, col="black")

In Figure 1 we can also see these general pattern of a higher effect with phylogenetic distance with the different regressions lines per family. I have done simple linear models, probably I should consider other alternatives that fit better the cloud of points. In Figure 1 the two coulored circles are there for a reason, once we consider just the crosses between the "far related species" we see a general pattern for solanaceae with almost all the values above 0.5. Solanaceae curiously belongs to the group with gametophytic incompatibility which is predicted to suffer more from heterospecific pollen. Moreover, for Brassicaceae (red points) again for the far related crosses we find kind of two groups with high effect and no effect. The no effect correspond to the species without incompatibility (over 50% of selfing) and the high effect to the species with strong incompatibility (0% of selfing), in order to show this in a better way I'm going just to plot these points of interest with their corresponding labels.
